---
title: "Features of APCalign"
author: "Fonti Kar"
date: "2023-10-05"
output: html_document
---

Other tools: 

- TNRS
- taxize
- ALA

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(TNRS, taxize, galah, APCalign, tidyverse, kableExtra, here)
```

### A subset of taxon strings to test across all tools

```{r}
# species <- read_csv(here("inst/extdata/species.csv"))
# species
# 
# consistency <- read_csv(here("tests/testthat/benchmarks/consistency_lookup.csv"))
# consistency

testing_taxa <- tibble(taxon = c("Banksia integrifolia", "Acacia aneura", "Commersonia rosea", "Grevillea umbellulata subsp. acerosa", "Fabaceae sp.", "Dryandra sp.", "Athrotaxis laxiflolia", "Galactia striata", "Hibbertia stricta", "Ecalypha indica australis"))
```

### {APCalign}

```{r}
resources <- load_taxonomic_resources("stable")

apcalign_output <- create_taxonomic_update_lookup(testing_taxa$taxon, resources = resources)

apcalign_output_full <- create_taxonomic_update_lookup(testing_taxa$taxon, resources = resources, full = TRUE)

apcalign_output
```

### {TNRS}

R package wrapper around the TNRS API. Web application found [here](https://tnrs.biendata.org/). Source code for [R package](https://github.com/EnquistLab/RTNRS)

```{r}
tnrs_output <- TNRS(testing_taxa$taxon)

tnrs_output
```

### {taxize}

```{r}
resolve_gnr_output <- resolve(testing_taxa$taxon, db = "gnr")

resolve_gnr_output |>
  as_tibble() |> 
  print(n = Inf)
```

### {galah}

Not exactly a taxon name resolver but users downloading biodiversity data would typically then use the ALA taxonomy 

```{r}
galah_config(atlas = "Australia")

galah_output <- search_taxa(testing_taxa |> rename(scientificName = taxon))

galah_output
```


## Make a features table

- Batch processing (Handles multiple names all at once)
- Fuzzy Matching
- Adjustments to Fuzzy matching
- Updates synonyms
- User Interface
- API
- Handles infraspecific taxa
- Handles taxonomic splits 
- Handles higher taxonomy
- Returns alternative taxa


TNRS does taxon rank, 
adding some things - browser interface; handles phrase names; identifies taxon rank of name; (adds identifiers — as in our `identifier` parameter; TNRS changes `Acacia sp.` to `Acacia`.

```{r}
options(knitr.kable.NA = '')
features <- data.frame(`Batch processing` = "X",
                       `Aligns taxonomic names` = "X",
                       `Updates taxonomic names` = c("X",NA, "X", NA),
                       `Fuzzy matching` = c("X", "X", "X", NA),
                       `Adjustable fuzzy matching` = c("X", rep(NA, 3)),
                       `Handles synonyms` = "X",
                       `Handles phrase names` = c("X", rep(NA, 3)),
                       `Web browser interface` = c("X", NA, "X", "X"),
                       `Identifies taxon rank of name` = c("X"),
                       `Handles infraspecific taxa` = c("X", NA, "X", "X"),
                       `Handles higher taxonomy` = "X",
                       `Handles taxonomic splits` = c("X", rep(NA, 3)),
                       check.names = FALSE) 

rownames(features) <- c("APCalign", "taxize", "TNRS", "kewr")

features |> 
  kable(align = "c") |> 
  kable_classic() |> 
  as_image(width = 4, file = here("output/table1_features.png"),
           zoom = 2)


  # add_header_above(c(" " = 1))
```

